---
output: github_document
---

<!-- toc: TRUE -->
<!-- toc_depth: 2 -->

<!-- README.md is generated from README.Rmd. Please edit that file -->

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end --> 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(tidyverse)
```


# {ggchalkboard} 

ggchalkboard is an example of a ggplot2 extension theme, which works on strategies to get to complementary geom/stat defaults `update_geom_defaults`!  The future looks a bit more coordinated!

I do not consider myself very gifted when it comes to thematic choices in ggplot2.  But thematic choices can be 'make or break' when it comes to audience.  And I do like chalkboards and think I can do a reasonable job mimicking their look.

This in-progress theme extension package tries to capture some best practices for creating theme extensions.  Unfortunately, due to my perceived lack of talent for theming, I'm actually pretty new to this space, so welcome feedback from anybody that has some real expertise to lend.

What might be new and interesting is the attempted coordination between aesthetic defaults for geom and stat layers and theme.  Below, we see for example that geom_point(), being which is defined by the GeomPoint proto object has the following default aesthetics.  

```{r}
ggplot2::GeomPoint$default_aes
```

These are sensible starting points.  Surely people would question default color choice of 'burlywood4'.  But 'black' color and 1.5 size defaults may look less sensible in the context of a great theme that you've constructed.  So we are going to tackle that with `update_geom_defaults()` can be used to complement a new theme function.  

But let's not get ahead of ourselves.  Let's start creating by creating theme_chalkboard, which is likely to be familiar as it is a ggplot2 extension entry point.  The topic gets first billing in the ggplot2 extension chapter in ggplot2 and is covered in some detail there, so we won't  https://ggplot2-book.org/extensions 

We'll be modifying a theme, since we don't want to start from scratch -- a theme contains a lot of decisions; to get a sense of that we can look at the length of the list object returned by `ggplot2::theme_gray()`

```{r}
ggplot2::theme_gray() |> length()
```

So let's get to writing our theme, `theme_chalkboard`.  

```{r theme_chalkboard}
#' Title
#'
#' @param board_color a string which is a color value
#' @param chalk_color a string which is a color value
#' @param ... 
#'
#' @return a ggplot2 object
#' @export
#'
#' @examples
theme_chalkboard <- function(board_color = "darkseagreen4",
                             chalk_color = "lightyellow", ...){

  ggplot2::theme_gray(...) %+replace%   ##<< we'll piggy back on an existing theme
    ggplot2::theme(
      rect = ggplot2::element_rect(fill = board_color,
                                   color = board_color),
      text = ggplot2::element_text(color = chalk_color,
                                   face = "italic",
                                   size = 18),
      panel.background = ggplot2::element_rect(fill = board_color,
                                               color = board_color),
      axis.text = ggplot2::element_text(color = chalk_color),
      axis.ticks = ggplot2::element_line(color = chalk_color),
      panel.grid = ggplot2::element_blank(),
      complete = TRUE   ##<< important, see 20.1.2 Complete themes in ggplot2 book
      )

}
```

*I think there is likely room for improvement here.  Please chime in if you'd like to alert me to problems and better ways to get things done!  I especially have hesitations about sizing and relative sizing.  Please excuse my ignorances.*

Let's try this out!  

```{r}
library(ggplot2)
ggplot(cars) +
  aes(speed, dist) +
  geom_point() 

last_plot() +
  theme_chalkboard()
```

This looks pretty good to me!  But ... who used a sharpie on the chalkboard!? Layers (geoms_* and stats_*) should be chalk-like too…

## Using `update_geom_defaults` for matching geom/stat layers look and feel of to new theme

We’ll use the ‘update_geom_defaults’ function to create a geoms_chalk_on() function, but before we do that, we anticipate wanting to turn that function back off. To do that, we’ll want to save ggplot2’s defaults. We save these lists as data in our theme package. We’ll use these to create geom_chalk_off().

Someday, this will be easier… But for now… https://github.com/tidyverse/ggplot2/pull/5098  

**This whole strategy might get immediately revised by copying the dev version of 5098 and using that here**

*Point of interest: the layer-defaults and theme coordination has been on wish lists for a long time: See https://github.com/tidyverse/ggplot2/pull/2749 'Allow default geom aesthetics to be set from theme'*

*This might be the precursor issue: https://github.com/tidyverse/ggplot2/issues/2239*


```{r DATASET, message=F}
## save default lists in a fresh session so that defaults are true to ggplot package defaults
default_aes_point   <- ggplot2::GeomPoint$default_aes
default_aes_segment <- ggplot2::GeomSegment$default_aes
default_aes_rug     <- ggplot2::GeomRug$default_aes
default_aes_rect    <- ggplot2::GeomRect$default_aes
default_aes_bar     <- ggplot2::GeomBar$default_aes
default_aes_label   <- ggplot2::GeomLabel$default_aes

# https://r-pkgs.org/data.htmlf
usethis::use_data(default_aes_point, overwrite = TRUE)
usethis::use_data(default_aes_segment, overwrite = TRUE)
usethis::use_data(default_aes_rug, overwrite = TRUE)
usethis::use_data(default_aes_rect, overwrite = TRUE)
usethis::use_data(default_aes_bar, overwrite = TRUE)
usethis::use_data(default_aes_label, overwrite = TRUE)

# A more general approach
# library("ggplot2")
# library(magrittr)
# 
# Geoms <- ls(pattern = '^Geom', env = as.environment('package:ggplot2'))
# 
# Grab all the default aes for geoms and assign them to value.
# for(i in 1:length(Geoms)){
#   
#   try(assign(paste0("default_aes", Geoms[i]), get(Geoms[i]) %>% .$default_aes))
#   
# }
```

```{r geoms_chalk_on}
geoms_chalk_on <- function(chalk_color = "lightyellow", color = NULL, fill = NULL){

  if(is.null(color)){color <- chalk_color}
  if(is.null(fill)){fill <- chalk_color}

  ggplot2::update_geom_defaults("point",   list(colour = color, size = 3.5,    alpha = .75))
  ggplot2::update_geom_defaults("segment", list(colour = color, size = 1.25,   alpha = .75))
  ggplot2::update_geom_defaults("rug",     list(colour = color, size = 1,      alpha = .75))
  ggplot2::update_geom_defaults("rect",    list(colour = color, fill = color, size = 1,      alpha = .75))
  ggplot2::update_geom_defaults("label",   list(fill = fill, color = "grey35", size = 5))
  

  # above is pretty limited... think more generally...  
  # https://stackoverflow.com/questions/21174625/ggplot-how-to-set-default-color-for-all-geoms
  # params <- ls(pattern = '^geom_', env = as.environment('package:ggplot2'))
  # geoms <- gsub("geom_", "", params)
  #
  # lapply(geoms, update_geom_defaults, list(colour = color))
  # lapply(geoms, update_geom_defaults, list(fill = fill))

}
```



```{r geoms_chalk_off}
geoms_chalk_off <- function(){

  ggplot2::update_geom_defaults("point",     default_aes_point)
  ggplot2::update_geom_defaults("segment",   default_aes_segment)
  ggplot2::update_geom_defaults("rug",       default_aes_rug)
  ggplot2::update_geom_defaults("rect",      default_aes_rect)
  ggplot2::update_geom_defaults("label",     default_aes_label)

}

```


```{r, echo = F, eval = F}
cache_geom_defaults <- new.env(parent = emptyenv())
cache_stats_defaults <- new.env(parent = emptyenv())

#' Modify geom/stat aesthetic defaults for future plots
#'
#' @param stat,geom Name of geom/stat to modify (like `"point"` or `"bin"`),
#'   or a Geom/Stat object (like `GeomPoint` or `StatBin`).
#' @param new Named list of aesthetics. Alternatively, `NULL` to reset the
#'   defaults.
#' @keywords internal
#' @name update_defaults
#' @export
#' @examples
#' # updating a geom's default aesthetic settings
#' # example: change geom_point()'s default color
#' GeomPoint$default_aes
#' update_geom_defaults("point", aes(color = "red"))
#' GeomPoint$default_aes
#' ggplot(mtcars, aes(mpg, wt)) + geom_point()
#'
#' # Reset defaults by using `new = NULL`
#' update_geom_defaults("point", NULL)
#'
#' # updating a stat's default aesthetic settings
#' # example: change stat_bin()'s default y-axis to the density scale
	@@ -29,27 +31,63 @@
#'   geom_function(fun = dnorm, color = "red")
#'
#' # reset default
#' update_stat_defaults("bin", NULL)

#' @rdname update_defaults
dev_update_geom_defaults <- function(geom, new) {
  if (is.null(new)) {

    vec_assert(geom, character(), 1)
    old <- cache_geom_defaults[[geom]]
    if (!is.null(old)) {
      new <- update_geom_defaults(geom, old)
      env_unbind(cache_geom_defaults, geom)
    }
    invisible(new)

  } else {

    g <- check_subclass(geom, "Geom", env = parent.frame())
    old <- g$default_aes
    # Only update cache the first time
    if (!geom %in% ls(cache_geom_defaults)) {
      cache_geom_defaults[[geom]] <- old
    }
    new <- rename_aes(new)
    new_names_order <- unique(c(names(old), names(new)))
    new <- defaults(new, old)[new_names_order]
    g$default_aes[names(new)] <- new
    invisible(old)

  }
}



```

**Another big revision is to work on generalizing this.  I know I've just selected a handful of important Geoms, but soon enough, we'll see mismatching**

Okay! Let's see our geoms_chalk_on in action.

```{r}
geoms_chalk_on()

last_plot()
```

Very nice! The points match the look and feel of the theme a lot better.  Okay, and let's check that we can turn the new defaults back off. 

```{r}
geoms_chalk_off()

last_plot() 
```

Great, it's working! 

While we are here, let's write a couple of derivative functions for a slightly different chalkboard look.

```{r theme_chalkboard_slate}
theme_chalkboard_slate <- function(...){

  theme_chalkboard(board_color = "lightskyblue4", chalk_color = "honeydew", ...)

}

geoms_chalk_on_slate <- function(chalk_color = "honeydew", ...){
  
  geoms_chalk_on(chalk_color = chalk_color, ...)
  
}
```


```{r}
geoms_chalk_on_slate()

last_plot() + 
  theme_chalkboard_slate() + 
  ggxmean:::geom_lm() + 
  ggxmean:::geom_lm_residuals()
```


```{r, eval = F, echo = F}
geoms_chalk_on()

readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2024/2024-02-20/isc_grants.csv") |>
  ggplot() + 
  aes(funded) +
  geom_histogram() + 
  theme_chalkboard()

```

Further coordination can be done when it comes to scales:

```{r}
scale_size_chalkboard <- function(...){
  
  scale_size(range = c(4, 12), ...)
  
}

geoms_chalk_on()

ggplot(data = cars) + 
  aes(x = speed, y = dist, size = dist) + 
  geom_point() + 
  theme_chalkboard() + 
  scale_size_chalkboard()
```

Color and fill scale are probably of greater interest, I know. Something to come back to. 

# Part II. Packaging and documentation  🚧

## Phase 1. Minimal working package

### Bit A. Created package archetecture, running `devtools::create(".")` in interactive session. ✅


```{r, eval = F}
devtools::create(".")
```

### Bit B. Added roxygen skeleton? 🚧 

Use a roxygen skeleton for auto documentation and making sure proposed functions are *exported*.  Generally, early on, I don't do much (anything) in terms of filling in the skeleton for documentation, because things may change.

### Bit C. Managed dependencies ? ✅

Package dependencies managed, i.e. `depend::function()` in proposed functions and declared in the DESCRIPTION

```{r, eval = F}
usethis::use_package("ggplot2")
```


### Bit D. Moved functions R folder? ✅

Use new {readme2pkg} function to do this from readme...

```{r}
library(tidyverse)
readme2pkg::chunk_to_r("theme_chalkboard")
readme2pkg::chunk_to_r("theme_chalkboard_slate")
readme2pkg::chunk_to_r("geoms_chalk_on")
readme2pkg::chunk_to_r("geoms_chalk_off")
readme2pkg::chunk_to_dir("DATASET", dir = "data-raw/")
```

### Bit E. Run `devtools::check()` and addressed errors. 🚧 ✅

```{r, eval = F}
devtools::check(pkg = ".")
```

### Bit F. Build package 🚧 ✅

```{r, eval = F}
devtools::build()
```


### Bit G. Write traditional README that uses built package (also serves as a test of build. ✅ 

The goal of the {ggchalkboard} package is to make it easy to theme ggplots like chalkboards

Install package with:

```
remotes::install_github("EvaMaeRey/ggchalkboard")
```

Once functions are exported you can remove go to two colons, and when things are are really finalized, then go without colons (and rearrange your readme...)

```{r, eval = F}
library(ggchalkboard)  
library(ggplot2)

ggplot(data = cars) + 
  aes(x = speed) + 
  geom_histogram() + 
  ggchalkboard:::theme_chalkboard()

ggchalkboard:::geoms_chalk_on()

last_plot()


ggchalkboard:::geoms_chalk_off()

last_plot()
```




### Bit H. Chosen a license? ✅


```{r, eval = F}
usethis::use_mit_license()
```

### Bit I. Add lifecycle badge (experimental) ✅

```{r, eval = F}
usethis::use_lifecycle_badge("experimental")
```


## Phase 2: Listen & iterate 🚧 

Try to get feedback from experts on API, implementation, default decisions.  Is there already work that solves this problem? 


## Phase 3: Let things settle

### Bit A. Settle on examples.  Put them in the roxygen skeleton and readme. 🚧 

### Bit B. Written formal tests of functions and save to test that folders 🚧 


That would look like this...

```{r test_calc_times_two_works, eval = F}
library(testthat)

test_that("calc times 2 works", {
  expect_equal(times_two(4), 8)
  expect_equal(times_two(5), 10)
  
})
```


```{r, eval = F}
readme2pkg::chunk_to_tests_testthat("test_calc_times_two_works")
```


### Bit C. Added a description and author information in the DESCRIPTION file 🚧 

### Bit D. Addressed *all* notes, warnings and errors. 🚧 

## Phase 4. Promote to wider audience...

### Bit A. Package website built?  ✅

```{r, eval = F}
usethis::use_pkgdown()
pkgdown::build_site()
```


### Bit B. Package website deployed? 🚧 ✅

## Phase 5: Harden/commit

### Submit to CRAN/RUniverse? 🚧 

# Appendix: Reports, Environment

## Edit Description file

```{r, eval = T}
readLines("DESCRIPTION")
```

## Environment

Here I just want to print the packages and the versions

```{r}
all <- sessionInfo() |> print() |> capture.output()
all[11:20]
```

## `devtools::check()` report

```{r, error = T, results="hide", warning=F, eval = F}
devtools::check(pkg = ".")
```
